---
title: "Tidy Tuesday - Week 9 (2021)"
author: "Julia Tache"
date: "2/23/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Employed Status 

This week's data are on employment and are sourced from the Bureau of Labor Statistics. I was interested this week in practicing my skills in plotly to create interactive bubble charts and investigating racial and sex wage gaps.  

The data show interesting trends, most notably that income inequity gaps seem to have been increasing overall and that the largest income gaps are between Asian men and women & White and Black men. 

```{r load data}
library(tidyverse)
tuesdata <- tidytuesdayR::tt_load('2021-02-23')
```

```{r aggregate data I}
library(reshape2)

earn <- tuesdata$earn
employed <- tuesdata$employed

agg <- earn %>%
  group_by(year, sex) %>%
  summarize(mean_yearly_earn = (mean(median_weekly_earn)*52)) %>% # approx. yearly income 
  filter(sex != "Both Sexes")

# reshape the data 
agg <- dcast(agg, year ~ sex)

# calculate gap between white and black earners 
agg <- agg %>%
  mutate(gap = (Men - Women)/100) 
# dollar amount in hundreds helps keep the graphs neat 

agg$Men <- agg$Men / 100
agg$Women <- agg$Women / 100
```

```{r Figure I}
library(plotly)

fig <- plot_ly(agg, x = ~Women, y = ~Men, text = ~year, type = 'scatter', mode = 'markers', marker = list(size = ~gap, color = ~gap, opacity = 0.5, colorbar = list(title = 'Income Gap'), colorscale='Viridis'))

fig %>% layout(title = 'Yearly Earning Gaps between All Men and Women (in hundreds of dollars)')
```

```{r aggregate data II}
agg2 <- earn %>%
  group_by(year, race, sex) %>%
  summarize(mean_yearly_earn = (mean(median_weekly_earn)*52)) %>%
  filter(sex != "Both Sexes") %>%
  filter(race != "All Races")

agg2$mean_yearly_earn <- agg2$mean_yearly_earn / 1000
```

```{r Figure II}
plot_ly(agg2, x = ~year, y = ~mean_yearly_earn, text = ~sex, type = 'scatter', mode = 'markers', color = ~race, marker = list(size = ~mean_yearly_earn, opacity = 0.5)) %>% layout(title = 'Average Yearly Earnings by Race and Sex (in thousands of dollars)')
```

```{r Adjust data for graphing}
agg2$mean_yearly_earn <- agg2$mean_yearly_earn * 10
agg2 <- dcast(agg2, year ~ race + sex)

agg2 <- agg2 %>% 
  mutate(asian_gap = Asian_Men - Asian_Women) %>%
  mutate(white_gap = White_Men - White_Women) %>%
  mutate(black_gap = `Black or African American_Men` - `Black or African American_Women`) %>%
  mutate(bw_men_gap = White_Men - `Black or African American_Men`) %>%
  mutate(bw_women_gap = White_Women - `Black or African American_Women`) 
```

```{r Figure III}
plot_ly(agg2, x = ~Asian_Women, y = ~Asian_Men, text = ~year, type = 'scatter', mode = 'markers', marker = list(size = ~asian_gap, opacity = 0.5, color = ~asian_gap, colorbar = list(title = 'Income Gap'), colorscale='Viridis')) %>% layout(title = 'Yearly Earning Gaps between Asian Men and Women (in hundreds of dollars)')
```
```{r Figure IV}
plot_ly(agg2, x = ~White_Women, y = ~White_Men, text = ~year, type = 'scatter', mode = 'markers', marker = list(size = ~white_gap, opacity = 0.5, color = ~white_gap, colorbar = list(title = 'Income Gap'), colorscale='Viridis')) %>% layout(title = 'Yearly Earning Gaps between White Men and Women (in hundreds of dollars)')
```
```{r Figure V}
plot_ly(agg2, x = ~`Black or African American_Men`, y = ~`Black or African American_Women`, text = ~year, type = 'scatter', mode = 'markers', marker = list(size = ~black_gap, opacity = 0.5, color = ~black_gap, colorbar = list(title = 'Income Gap'), colorscale='Viridis')) %>% layout(title = 'Yearly Earning Gaps between Black Men and Women (in hundreds of dollars)')
```


```{r Figure VI}
plot_ly(agg2, x = ~`White_Men`, y = ~`Black or African American_Men`, text = ~year, type = 'scatter', mode = 'markers', marker = list(size = ~bw_men_gap, opacity = 0.5, color = ~bw_men_gap, colorbar = list(title = 'Income Gap'), colorscale='Viridis')) %>% layout(title = 'Yearly Earning Gaps between Black and White Men (in hundreds of dollars)')
```

```{r Figure VII}
plot_ly(agg2, x = ~`White_Women`, y = ~`Black or African American_Women`, text = ~year, type = 'scatter', mode = 'markers', marker = list(size = ~bw_women_gap, opacity = 0.5, color = ~bw_women_gap, colorbar = list(title = 'Income Gap'), colorscale='Viridis')) %>% layout(title = 'Yearly Earning Gaps between Black and White Women (in hundreds of dollars)')
```


