---
title: "Tidy Tuesday - Week 3"
author: "Julia Tache"
date: "1/18/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


```{r load data}
library(tidyverse)
library(tidytuesdayR)

tuesdata <- tidytuesdayR::tt_load('2022-01-18')
chocolate <- tuesdata$chocolate

summary(chocolate)
head(chocolate)
```

```{r average ratings}
library(countrycode)

chocolate$cocoa_percent <- as.numeric(str_extract(chocolate$cocoa_percent, "\\d+"))/100

dark_chocolate <- chocolate %>%
  filter(cocoa_percent <= 0.8) %>%
  filter(country_of_bean_origin != "Blend") %>%
  group_by(review_date, country_of_bean_origin) %>%
  summarize(country_ratings = mean(rating))

extra_dark_chocolate <- chocolate %>%
  filter(cocoa_percent > 0.8) %>%
  filter(country_of_bean_origin != "Blend") %>%
  group_by(review_date, country_of_bean_origin) %>%
  summarize(country_ratings = mean(rating)) 

dark_chocolate$continent <- countrycode(sourcevar = dark_chocolate$country_of_bean_origin,
                                        origin = "country.name",
                                        destination = "continent")

extra_dark_chocolate$continent <- countrycode(sourcevar = extra_dark_chocolate$country_of_bean_origin,
                                        origin = "country.name",
                                        destination = "continent")

dark_chocolate <- na.omit(dark_chocolate)
extra_dark_chocolate <- na.omit(extra_dark_chocolate)

extra_dark_chocolate <- extra_dark_chocolate %>%
  filter(continent != "Asia")

head(dark_chocolate)
head(extra_dark_chocolate)
```


```{r top words}
library(tokenizers)

dark_choc <- chocolate %>%
  filter(cocoa_percent <= 0.8)

dark_choc_words <- strsplit(toString(dark_choc$most_memorable_characteristics),",")
dark_choc_words_freq <- table(unlist(dark_choc_words))
dark_choc_words_df <- as.data.frame(cbind(names(dark_choc_words_freq), as.numeric(dark_choc_words_freq)))
dark_choc_words_df$V2 <- as.numeric(dark_choc_words_df$V2)

xtra_dark_choc <- chocolate %>%
  filter(cocoa_percent > 0.8)

xtra_dark_choc_words <- strsplit(toString(xtra_dark_choc$most_memorable_characteristics),",")
xtra_dark_choc_words_freq <- table(unlist(xtra_dark_choc_words))
xtra_dark_choc_words_df <- as.data.frame(cbind(names(xtra_dark_choc_words_freq), as.numeric(xtra_dark_choc_words_freq)))
xtra_dark_choc_words_df$V2 <- as.numeric(xtra_dark_choc_words_df$V2)
```

```{r dark chocolate ridge graph}
library(ggridges)
library(wesanderson)
library(hrbrthemes)

pal <- wes_palette("GrandBudapest1", 100, type = "continuous")

ggplot(dark_chocolate, aes(x = country_ratings, y = continent, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 1, rel_min_height = 0.01) +
  scale_fill_gradientn(colours = pal, guide = guide_colourbar(), name = "Rating Scale") +
  labs(title = "Dark Chocolate Review Ratings by Continent (2006-2021)") +
  theme_ridges() + 
  theme(plot.title = element_text(face = "italic", hjust = 0.5),
        legend.position = "bottom",
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 7),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 12),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) + 
  guides(fill = guide_colourbar(label.position = "bottom", barwidth = 25, barheight = 0.5))

ggsave("ridgeplot_1.png")
```

This first graph is a density ridgeplot that takes a look at the distribution of ratings for dark chocolate (cocoa <=80%) amongst continents where the chocolate in the data were sourced from (Oceania, Asia, the Americas, and Africa). Most chocolates peak between a score of 3 and a 3.5 on a 5-point scale. Both Africa and the Americas have long tails on the left indicating a collection of chocolates with lower ratings. The continuous color scale is "GrandBudapest1" from the Wes Anderson package and goes from orange to pink to maroon to mauve.

```{r extra dark chocolate ridge graph}
ggplot(extra_dark_chocolate, aes(x = country_ratings, y = continent, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 0.8, rel_min_height = 0.01) +
  scale_fill_gradientn(colours = pal, guide = guide_colourbar(), name = "Rating Scale") +
  labs(title = "Extra Dark Chocolate Review Ratings by Continent (2006-2021)") +
  theme_ridges() + 
  theme(plot.title = element_text(face = "italic", hjust = 0.5),
        legend.position = "bottom",
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 7),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 12),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) + 
  guides(fill = guide_colourbar(label.position = "bottom", barwidth = 25, barheight = 0.5))

ggsave("ridgeplot_2.png")
```

This graph is another density ridgeplot that takes a look at the distribution of ratings for extra dark chocolate (cocoa >80%) from the different continents in the data. Both Oceania and Africa have bimodal distrubtions which peak between a score of 2 and 3 and a score of 3 and 4. The Americans has a long tail to the left and a peak right after a score of 3. The color scale is once again "GrandBudapest1" from the Wes Anderson package.

```{r dark chocolate word cloud}
library(webshot)
webshot::install_phantomjs()

library(wordcloud2) 

wc1 <- wordcloud2(data = dark_choc_words_df, size = 1, color= "random-dark")
wc1

wc2 <- wordcloud2(data = xtra_dark_choc_words_df, size = 1, color= "random-dark")
wc2

library(htmlwidgets)
saveWidget(wc1,"dark-chocolate-cloud.html", selfcontained = F)
webshot("dark-chocolate-cloud.html","dark-chocolate-cloud.png", delay = 5, vwidth = 480, vheight = 480)

saveWidget(wc2,"extra-dark-chocolate-cloud.html", selfcontained = F)
webshot("extra-dark-chocolate-cloud.html","extra-dark-chocolate-cloud.png", delay = 5, vwidth = 480, vheight = 480)
```

The first multicolored wordcloud gives the tasting notes of dark chocolate. Some of the top notes are cocoa, creamy, earthy, intense, and floral. 

The second multicolored wordcloud gives the tasting notes of extra dark chocolate. Some of the top notes are bitter, intense, fatty, and spicy. 
